High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-26 notebooks/eccco.qmd
some work on illustrative example

>>>>

```{julia}
# Counteractual data:
counterfactual_data = load_circles(2000)
X = counterfactual_data.X
y = counterfactual_data.y
labels = counterfactual_data.output_encoder.labels
input_dim, nobs = size(X)
batch_size = Int(round(nobs/10))
```

```{julia}
𝒟x = Normal()
𝒟y = Categorical(ones(2) ./ 2)
sampler = ConditionalSampler(𝒟x, 𝒟y, input_size=size(X)[1:end-1], batch_size=batch_size)
clf = JointEnergyClassifier(
    sampler;
    builder=MLJFlux.MLP(hidden=(32, 32, 32,), σ=Flux.relu),
    batch_size=batch_size,
    finaliser=x -> x,
    loss=Flux.Losses.logitcrossentropy,
    jem_training_params=(
        α=[1.0,1.0,1e-1],
        verbosity=10,
    ),
)
```

```{julia}
method = :simple_inductive
cov = .95
conf_model = conformal_model(clf; method=method, coverage=cov)
mach = machine(conf_model, table(permutedims(X)), labels)
fit!(mach)
```

```{julia}
#| echo: false

jem = mach.model.model.jem
batch_size = mach.model.model.batch_size
X = Float32.(matrix(X))
if typeof(jem.sampler) <: ConditionalSampler

    plts = []
    for target in 1:2
        X̂ = generate_conditional_samples(jem, batch_size, target; niter=200) 
        ex = extrema(hcat(X,X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=100)
        x2 = range(1.0f0.*ylims...,length=100)
        plt = Plots.contour(
            x1, x2, (x, y) -> softmax(jem([x, y]))[target], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(X[1,:], X[2,:], color=Int.(labels.refs), group=Int.(labels.refs), alpha=0.5)
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target], size(X̂,2)), 
            group=repeat([target], size(X̂,2)), 
            shape=:star5, ms=10
        )
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400))
    display(plt)
end
```

```{julia}
M = ECCCo.ConformalModel(mach.model, mach.fitresult)
factual_label = 0
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 1
factual = predict_label(M, counterfactual_data, x_factual)[1]

# ECCCo:
λ=[0.5,0.1,0.5]
temp=0.5
η=0.01

# Generate counterfactuals using ECCCo generator:
generator = ECCCoGenerator(
    λ=λ, 
    temp=temp, 
    opt=Flux.Optimise.Adam(η),
)

ce = CounterfactualExplanation(x_factual, target, counterfactual_data, M, generator)
```
